Scikit-learn: machine learning without learning the machinery

### 2014-07-01 article.tex
random forests in lowercase
Uppercase variant is used as a trademark.

% Introduce and describe the data
Machine learning is about extracting rules from data, most often with the
goal of making decision on new data \cite{elemstatlearn}. As a simple example, we show on
Fig.\,\ref{fig:data} data about wages in the US, from \cite{berndt1991}:
11 \emph{features}, including wage, number of years of education and of work
experience, on 534 individuals -- called \emph{samples} in machine-learning
jargon.
Code downloading the data to reproduce the figures can be found on
\url{https://github.com/scikit-learn/SigMobile-paper}.
%
This data showcases many of the typical difficulties in machine learning.
Samples are very irregularly distributed, as most individuals
finished studies after high-school. As a result, there are gaping
holes in the 2D plot of education versus work experience, in which any
statistical analysis is extrapolation. The data is very noisy: for a
given education and work experience wages vary widely. This variability
is probably explained by missing factors, such as sector of
activity, but adding them to the analysis creates a more complex picture,
in 3D or more, rather than 2D, with even more gaps.

On Fig.\,\ref{fig:random_forest}, we use a popular machine-learning
algorithm, Random Forests, to predict the wage from the years of
education, or the years of work experience, either as separate features,
or combined. The patchy square appearance of the prediction is due to the
inners of the algorithm, and it illustrates the corresponding
extrapolation mechanism. The predictions fit the data very; maybe too
well, as, on the prediction solely from the years of education (left of
the figure), it is hard to believe in the bump at 5 years. This bump is
probably a classic case of \emph{overfit}: the algorithm is learning its
prediction from noise in the data. As a result, the prediction error on
new data will be significantly different than that measured on the
training data.

>>>>

% Introduce and describe the data
Machine learning is about extracting rules from data, most often with the
goal of making decision on new data \cite{elemstatlearn}. As a simple example, we show on
Fig.\,\ref{fig:data} data about wages in the US, from \cite{berndt1991}:
11 \emph{features}, including wage, number of years of education and of work
experience, on 534 individuals -- called \emph{samples} in machine-learning
jargon.
Code downloading the data to reproduce the figures can be found on
\url{https://github.com/scikit-learn/SigMobile-paper}.
%
This data showcases many of the typical difficulties in machine learning.
Samples are very irregularly distributed, as most individuals
finished studies after high-school. As a result, there are gaping
holes in the 2D plot of education versus work experience, in which any
statistical analysis is extrapolation. The data is very noisy: for a
given pair of education and work experience values, wages vary widely.
This variability
is probably explained by missing factors, such as sector of
activity, but adding them to the analysis creates a more complex picture,
in 3D or more, rather than 2D, with even more gaps.

On Fig.\,\ref{fig:random_forest}, we use a popular machine-learning
algorithm, random forests, to predict the wage from the years of
education, or the years of work experience, either as separate features,
or combined. The patchy square appearance of the prediction is due to the
inners of the algorithm, and it illustrates the corresponding
extrapolation mechanism. The predictions fit the data very; maybe too
well, as, on the prediction solely from the years of education (left of
the figure), it is hard to believe in the bump at 5 years. This bump is
probably a classic case of \emph{overfit}: the algorithm is learning its
prediction from noise in the data. As a result, the prediction error on
new data will be significantly different than that measured on the
training data.
